The evolution of the transcriptome of mammalian placentas

### 2016-09-08 manuscript/manuscript.Rnw
remove commented out latex code
change title
genes are now in supplemental

%\documentclass[english,12pt]{article}
\documentclass[5p]{elsarticle_dla}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\setmathrm{FreeSerif}
\setmathsf{FreeSans}
\setmathtt{FreeSerif}
%\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage[letterpaper]{geometry}
%\usepackage{url}
\usepackage{graphicx}
%\usepackage{rotating}
%\usepackage{wrapfig}
\usepackage{booktabs}
%\usepackage{float}
\usepackage{multirow}
%\usepackage{setspace}
\usepackage{acronym}
\usepackage{array}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=numeric-comp,style=nature,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=black,citecolor=blue]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads]{endfloat}
% \renewcommand{\efloatseparator}{\mbox{}}
%\AtBeginFigures{\newpage \renewcommand{\efloatseparator}{\mbox{}}\section*{Figure Legends}}
%\AtBeginTables{\newpage \section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
% \renewcommand{\includegraphics}[2][]{}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\newcommand{\DEW}[1]{\textcolor{purple}{\fxnote{DEW: #1}}}
% \usepackage[x11names,svgnames]{xcolor}
% \usepackage{texshade}
% \usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
% this package will stretch the font slightly if necessary.
% \usepackage{microtype}
\usepackage{zref-xr,zref-user}
\zxrsetup{toltxlabel}
\zexternaldocument*{supplemental_material}
\usepackage[capitalize]{cleveref}
% \newenvironment{narrow}[2]{%
%   \begin{list}{}{%
%       \setlength{\topsep}{0pt}%
%       \setlength{\leftmargin}{#1}%
%       \setlength{\rightmargin}{#2}%
%       \setlength{\listparindent}{\parindent}%
%       \setlength{\itemindent}{\parindent}%
%       \setlength{\parsep}{\parskip}}%
%   \item[]}{\end{list}}
% \def\newblock{\hskip}
% \newenvironment{paperquote}{%
%   \begin{quote}%
%      \it
%   }%
%   {\end{quote}}
% \renewcommand{\textfraction}{0.15}
% \renewcommand{\topfraction}{0.85}
% \renewcommand{\bottomfraction}{0.65}
% \renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
% \newenvironment{enumerate*}%
%   {\begin{enumerate}%
%     \setlength{\itemsep}{0pt}%
%     \setlength{\parskip}{0pt}}%
%   {\end{enumerate}}
% \newenvironment{itemize*}%
%   {\begin{itemize}%
%     \setlength{\itemsep}{0pt}%
%     \setlength{\parskip}{0pt}}%
%   {\end{itemize}}
% \setcounter{tocdepth}{3}
% \oddsidemargin 0.0in 
% \textwidth 6in
% \raggedbottom
% \clubpenalty = 10000
% \widowpenalty = 10000
% set a large emergency stretch to avoid issues with overfull boxes
% \setlength{\emergencystretch}{20em}
% make the afill smaller
% \renewcommand\Affilfont{\itshape\small}
%\renewcommand\cite[1]{\autocite{#1}}
%\renewcommand\citet[1]{\textcite{#1}}
% \pagestyle{fancy}

@@

For the term placenta of \smonkey, assembled contigs from RNA-seq were
aligned using against \human and the best alignment was chosen as the
annotation for each assembled contig. In the term placenta of
\smonkey, the top 50 non-housekeeping genes
(\cref{tab:top50_spider_monkey_placenta}) include a possible regulator
of androgen receptor signaling
\cite{Sampson.ea2012:PAGE4positivityisassociated}, \textit{PAGE4} %
(\Sexpr{gene.expression("ateles_fusciceps","PAGE4")}), %
the pregnancy specific glycoprotein \textit{PSG8}
(\Sexpr{gene.expression("ateles_fusciceps","PSG8")}), %
only one member of the chorionic somatotropins, \textit{GH1} %
(\Sexpr{gene.expression("ateles_fusciceps","GH1")}), %
a few HLA genes (\textit{HLA-A}, \textit{HLA-F},
\cref{tab:top50_spider_monkey_placenta}), and multiple components of
annexin complexes
(\Sexpr{gene.expression("ateles_fusciceps","ANXA2",complete=TRUE)}, %
\Sexpr{gene.expression("ateles_fusciceps","S100P",complete=TRUE)}, %
\Sexpr{gene.expression("ateles_fusciceps","S100A11",complete=TRUE)}) %
which may play a role in angiogenesis, and certainly regulate
cell-cell interaction in the developing placenta
\cite{Liu.ea2015:AnnexinA2complexeswith}.
\Sexpr{gene.expression("ateles_fusciceps","AL365202.1",complete=TRUE)}
is also highly expressed, has no known function, and is duplicated
multiple times in \textit{Macaca mulatta} (Ensembl gene tree
\href{http://useast.ensembl.org/Multi/GeneTree/Image?gt=ENSGT00580000081348}{ENSGT00580000081348}).
All genes with expression greater than 50 \ac{FPKM} are listed in
Supplementary \cref{tab:all_ateles_fusciceps_placenta}.

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \mouse are shown in
\cref{tab:top50_mouse_placenta} with a list of all genes with
$\text{FPKM} \ge 50$ in Supplemental
\cref{tab:all_mus_musculus_placenta}. The top expressed genes include
\Sexpr{gene.expression("mus_musculus","Mir6236",complete=TRUE)}, whose
target \textit{in silico} is the orphan G-coupled receptor
\textit{Mrgprg}
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget},
\Sexpr{gene.expression("mus_musculus","H19",complete=TRUE)} and
\Sexpr{gene.expression("mus_musculus","Gm27786",complete=TRUE)}, both
of which are maternally imprinted (the later of which is a homolog of
human \textit{H19\_2}),
\Sexpr{gene.expression("mus_musculus","Igf2",complete=TRUE)}, numerous
cathepsins and a rodentia specific expansion of prolactin related
genes. There are also two miRNA which overlap with the of the 7S
signal recognition particle
(\Sexpr{gene.expression("mus_musculus","AC157822.1",complete=TRUE)}
and
\Sexpr{gene.expression("mus_musculus","AC099934.1",complete=TRUE)})
and numerous pseudogenes
(\Sexpr{gene.expression("mus_musculus","Gm11966",complete=TRUE)}
\Sexpr{gene.expression("mus_musculus","Gm10736",complete=TRUE)}
\Sexpr{gene.expression("mus_musculus","Gm10180",complete=TRUE)}
\Sexpr{gene.expression("mus_musculus","Gm29216",complete=TRUE)}) which
appear to be highly expressed, however, the short read length (34--44)
in many samples is likely preventing unambiguous assignment of all
reads to the more lengthy complete transcript, so these shorter
transcripts should be treated with suspicion, and we will ignore them
in further discussion. The prolactin-related genes
(\Sexpr{gene.expression("mus_musculus","Prl3b1",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Prl2b1",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Prl8a9",complete=TRUE)}, and
\Sexpr{gene.expression("mus_musculus","Prl8a6",complete=TRUE)}) are
very highly expressed in \mouse in, with %
\Sexpr{combined.fpkm[species=="mus musculus" &
  grepl("Prl(\\d.\\d)?",gene_short_name), sum(mean_fpkm >= 1)]}
prolactin related genes having $\text{FPKM} \ge 1$
(\cref{fig:prolactin_evolution_tree}, Supplemental
\cref{fig:prolactin_receptor_evolution_tree}). The cathespins have
also expanded in rodents, with six showing high expression
(\Sexpr{gene.expression("mus_musculus","Ctsj",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Cts6",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Ctsr",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Ctsq",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Cts3",complete=TRUE)}, and
\Sexpr{gene.expression("mus_musculus","Ctsm",complete=TRUE)},
\cref{tab:top50_mouse_placenta,fig:pag2_prss16_ctsl_psg_evolution_tree}).
Finally, \textit{Peg10}, an imprinted gene required for proper
formation of the placenta in mice is also highly expressed
(\Sexpr{gene.expression("mus_musculus","Peg10")}), as previously
reported \cite{Ono.ea2006:DeletionofPeg10imprinted}.

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \galili are shown in
\cref{tab:top50_n_galili_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_spalax_galili_placenta}.
Components of annexin complexes are highly expressed
(\Sexpr{gene.expression("spalax_galili","S100a6",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","S100a11",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","S100a10",complete=TRUE)},
and
\Sexpr{gene.expression("spalax_galili","Anxa2",complete=TRUE)}),
as are members of the thymosin family
(\Sexpr{gene.expression("spalax_galili","Tmsb4x",complete=TRUE)}
and
\Sexpr{gene.expression("spalax_galili","Tmsb10",complete=TRUE)}),
various cytoskeleton related proteins
(\Sexpr{gene.expression("spalax_galili","Actc1",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","Actg2",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","Acta2",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","Myl9",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","Vim",complete=TRUE)}), and
transmembrane protein of currently unknown function, 
\Sexpr{gene.expression("spalax_galili","Gm10134",complete=TRUE)}.

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \carmeli are shown in
\cref{tab:top50_s_carmeli_placenta} with a list of all highly
expressed genes in Supplemental
\cref{tab:all_spalax_carmeli_placenta}. Numerous hemoglobin subunits
ware highly expressed
(\Sexpr{gene.expression("spalax_carmeli","Hbb-bh1",complete=TRUE)},
\Sexpr{gene.expression("spalax_carmeli","Hba-a1",complete=TRUE)}, and
\Sexpr{gene.expression("spalax_carmeli","Hbb-bs",complete=TRUE)}), as
well as the relaxin
\Sexpr{gene.expression("spalax_carmeli","Rln1",complete=TRUE)},
multiple cathespins
(\Sexpr{gene.expression("spalax_carmeli","Ctsm",complete=TRUE)},
\Sexpr{gene.expression("spalax_carmeli","BC051665",complete=TRUE)},
and \Sexpr{gene.expression("spalax_carmeli","Ctsl",complete=TRUE)},
\cref{fig:pag2_prss16_ctsl_psg_evolution_tree}~B), and many prolactin-related genes
(\Sexpr{gene.expression("spalax_carmeli","Prl4a1",complete=TRUE)},
\Sexpr{gene.expression("spalax_carmeli","Prl6a1",complete=TRUE)},
\Sexpr{gene.expression("spalax_carmeli","Prl3b1",complete=TRUE)}, and
\Sexpr{gene.expression("spalax_carmeli","Prl7c1",complete=TRUE)},
\cref{fig:prolactin_evolution_tree}) which shows a similar pattern of
expansion seen in \mouse.

@@

The top 50 non-housekeeping expressed genes as measured by \ac{FPKM}
in the term placenta of \dog are shown in
\cref{tab:top50_dog_placenta} with a list of all genes in Supplemental
\cref{tab:all_canis_familiaris_placenta}. The gene with the highest
expression %
(\Sexpr{gene.expression("canis familiaris","TMSB10")}) is the thymosin 
\textit{TMSB10} %
(\Sexpr{gene.expression("canis familiaris","ENSCAFG00000002115")}) %
is the hormone relaxin (\textit{RLN2}). Other highly expressed genes
among the top 50 (\cref{tab:top50_dog_placenta}) include genes
involved in the cytoskeleton (\textit{TMSB10}, \textit{TMSB4X}),
anticoagulation (\textit{TFPI2}, \textit{PLA2G7}), innate immunity
(\textit{LCN2}, \textit{GM2A}), lipid metabolism (\textit{APOE}),
inhibition of matrix metalloproteinases (\textit{TIMP2},
\textit{TIMP3}), and growth factors (\textit{PGF}, \textit{EGFR},
\textit{HTRA1}). The gene \textit{LGALS1}, which shows evidence of
immunosupression during pregnancy
\cite{Than.ea2008:Emergenceofhormonaland}, is highly expressed %
(\Sexpr{gene.expression("canis familiaris","LGALS1")}). %
In addition, an ortholog of \textit{SSLP1} (ENSCAFG00000010638) a gene
expressed in the placentas of \cow (%
\Sexpr{gene.expression("bos taurus","SSLP1",complete=TRUE)}, %
\cref{tab:top50_cow_placenta}) and \sheep
(\Sexpr{gene.expression("ovis aries","SOLD1",complete=TRUE)},
\cref{tab:top50_sheep_placenta}) but is not present in primates
\cite{Ushizawa.ea2009:Characterizationandexpressionanalysis}
(\cref{fig:sslp1_evolution_tree}), is also expressed at high levels
%
(\Sexpr{gene.expression("canis familiaris","ENSCAFG00000010638")}). %
There is no placental expression of the estrogen synthase gene
(\textit{CYP19A1}) in \dog 
(\Sexpr{gene.expression("canis familiaris","CYP19A1")}
\cref{fig:estrogen_evolution_tree}), 
and expression of \textit{HSD3B7} is also low
(\Sexpr{gene.expression("canis familiaris","HSD3B7")}, 
\cref{fig:progesterone_evolution_tree}).

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \textit{Sus scrofa} are shown in
\cref{tab:top50_pig_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_sus_scrofa_placenta}. The top two
expressed transcripts,
\Sexpr{gene.expression("sus_scrofa","ENSSSCG00000025162",complete=TRUE)}
and
\Sexpr{gene.expression("sus_scrofa","ENSSSCG00000021356",complete=TRUE)},
are miRNA which correspond to the \cow miRNA
\href{http://www.mirbase.org/cgi-bin/mirna_entry.pl?acc=MI0013080}{bta-mir-2904-1}
and
\href{http://www.mirbase.org/cgi-bin/mirna_entry.pl?acc=MIMAT0013845}{bta-mir-2887-1},
respectively, both of which are also expressed in \cow ovary
\cite{Hossain.ea2009:Identificationandcharacterizationof}, although
neither of which have an obvious target gene on the basis of
\textit{in silico} analysis
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget}. Other highly
expressed genes include members of the insulin growth factor family
(\Sexpr{gene.expression("sus_scrofa","IGF2",complete=TRUE)} and
\Sexpr{gene.expression("sus_scrofa","IGFBP3",complete=TRUE)}), members
of annexin complexes
(\Sexpr{gene.expression("sus_scrofa","S100A1",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","ANXA2",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","S100A12",complete=TRUE)}, and
\Sexpr{gene.expression("sus_scrofa","S100A10",complete=TRUE)}), the
galectin \Sexpr{gene.expression("sus_scrofa","LGALS3",complete=TRUE)}
and multiple cytoskeleton and extracellular matrix proteins
(\Sexpr{gene.expression("sus_scrofa","KRT19",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","COL1A1",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","BETA-ACTIN",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","KRT9",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","CLDN4",complete=TRUE)}, and
\Sexpr{gene.expression("sus_scrofa","KRT9",complete=TRUE)}).

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \horse are shown in
\cref{tab:top50_horse_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_equus_caballus_placenta}. The gene
with the highest expression was the relaxin,
\Sexpr{gene.expression("equus_caballus","RLN",complete=TRUE)},
followed by two miRNA
(\Sexpr{gene.expression("equus_caballus","ENSECAG00000025328",complete=TRUE)}
and
\Sexpr{gene.expression("equus_caballus","ENSECAG00000027056",complete=TRUE)})
with no known \textit{in silico} targets
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget}, the thymosin
\Sexpr{gene.expression("equus_caballus","TMSB4",complete=TRUE)},
members of the insulin growth factor family
(\Sexpr{gene.expression("equus_caballus","IGF2",complete=TRUE)} and
\Sexpr{gene.expression("equus_caballus","IGFBP3",complete=TRUE)}) and
multiple members of annexin complexes
(\Sexpr{gene.expression("equus_caballus","S100A6",complete=TRUE)},
\Sexpr{gene.expression("equus_caballus","S100A1",complete=TRUE)}, and
\Sexpr{gene.expression("equus_caballus","ANXA2",complete=TRUE)}),
similar to the pattern of expression seen in \pig and \dog.

>>>>

%\documentclass[english,12pt]{article}
\documentclass[5p]{elsarticle_dla}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\setmathrm{FreeSerif}
\setmathsf{FreeSans}
\setmathtt{FreeSerif}
%\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage[letterpaper]{geometry}
%\usepackage{url}
\usepackage{graphicx}
%\usepackage{rotating}
%\usepackage{wrapfig}
\usepackage{booktabs}
%\usepackage{float}
\usepackage{multirow}
%\usepackage{setspace}
\usepackage{acronym}
\usepackage{array}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=numeric-comp,style=nature,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=black,citecolor=blue]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads]{endfloat}
% \renewcommand{\efloatseparator}{\mbox{}}
%\AtBeginFigures{\newpage \renewcommand{\efloatseparator}{\mbox{}}\section*{Figure Legends}}
%\AtBeginTables{\newpage \section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
% \renewcommand{\includegraphics}[2][]{}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\newcommand{\DEW}[1]{\textcolor{purple}{\fxnote{DEW: #1}}}
% \usepackage[x11names,svgnames]{xcolor}
% \usepackage{texshade}
% \usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
% this package will stretch the font slightly if necessary.
% \usepackage{microtype}
\usepackage{zref-xr,zref-user}
\zxrsetup{toltxlabel}
\zexternaldocument*{supplemental_material}
\usepackage[capitalize]{cleveref}

@@

For the term placenta of \smonkey, assembled contigs from RNA-seq were
aligned using against \human and the best alignment was chosen as the
annotation for each assembled contig. In the term placenta of
\smonkey, the top 50 non-housekeeping genes
(Supplemental \cref{tab:all_spider_monkey_placenta}) include a possible regulator
of androgen receptor signaling
\cite{Sampson.ea2012:PAGE4positivityisassociated}, \textit{PAGE4} %
(\Sexpr{gene.expression("ateles_fusciceps","PAGE4")}), %
the pregnancy specific glycoprotein \textit{PSG8}
(\Sexpr{gene.expression("ateles_fusciceps","PSG8")}), %
only one member of the chorionic somatotropins, \textit{GH1} %
(\Sexpr{gene.expression("ateles_fusciceps","GH1")}), %
a few HLA genes (\textit{HLA-A}, \textit{HLA-F},
Supplemental \cref{tab:all_spider_monkey_placenta}), and multiple components of
annexin complexes
(\Sexpr{gene.expression("ateles_fusciceps","ANXA2",complete=TRUE)}, %
\Sexpr{gene.expression("ateles_fusciceps","S100P",complete=TRUE)}, %
\Sexpr{gene.expression("ateles_fusciceps","S100A11",complete=TRUE)}) %
which may play a role in angiogenesis, and certainly regulate
cell-cell interaction in the developing placenta
\cite{Liu.ea2015:AnnexinA2complexeswith}.
\Sexpr{gene.expression("ateles_fusciceps","AL365202.1",complete=TRUE)}
is also highly expressed, has no known function, and is duplicated
multiple times in \textit{Macaca mulatta} (Ensembl gene tree
\href{http://useast.ensembl.org/Multi/GeneTree/Image?gt=ENSGT00580000081348}{ENSGT00580000081348}).
All genes with expression greater than 50 \ac{FPKM} are listed in
Supplementary \cref{tab:all_ateles_fusciceps_placenta}.

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \mouse are shown in
Supplemental \cref{tab:all_mus_musculus_placenta} with a list of all genes with
$\text{FPKM} \ge 50$ in Supplemental
\cref{tab:all_mus_musculus_placenta}. The top expressed genes include
\Sexpr{gene.expression("mus_musculus","Mir6236",complete=TRUE)}, whose
target \textit{in silico} is the orphan G-coupled receptor
\textit{Mrgprg}
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget},
\Sexpr{gene.expression("mus_musculus","H19",complete=TRUE)} and
\Sexpr{gene.expression("mus_musculus","Gm27786",complete=TRUE)}, both
of which are maternally imprinted (the later of which is a homolog of
human \textit{H19\_2}),
\Sexpr{gene.expression("mus_musculus","Igf2",complete=TRUE)}, numerous
cathepsins and a rodentia specific expansion of prolactin related
genes. There are also two miRNA which overlap with the of the 7S
signal recognition particle
(\Sexpr{gene.expression("mus_musculus","AC157822.1",complete=TRUE)}
and
\Sexpr{gene.expression("mus_musculus","AC099934.1",complete=TRUE)})
and numerous pseudogenes
(\Sexpr{gene.expression("mus_musculus","Gm11966",complete=TRUE)}
\Sexpr{gene.expression("mus_musculus","Gm10736",complete=TRUE)}
\Sexpr{gene.expression("mus_musculus","Gm10180",complete=TRUE)}
\Sexpr{gene.expression("mus_musculus","Gm29216",complete=TRUE)}) which
appear to be highly expressed, however, the short read length (34--44)
in many samples is likely preventing unambiguous assignment of all
reads to the more lengthy complete transcript, so these shorter
transcripts should be treated with suspicion, and we will ignore them
in further discussion. The prolactin-related genes
(\Sexpr{gene.expression("mus_musculus","Prl3b1",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Prl2b1",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Prl8a9",complete=TRUE)}, and
\Sexpr{gene.expression("mus_musculus","Prl8a6",complete=TRUE)}) are
very highly expressed in \mouse in, with %
\Sexpr{combined.fpkm[species=="mus musculus" &
  grepl("Prl(\\d.\\d)?",gene_short_name), sum(mean_fpkm >= 1)]}
prolactin related genes having $\text{FPKM} \ge 1$
(\cref{fig:prolactin_evolution_tree}, Supplemental
\cref{fig:prolactin_receptor_evolution_tree}). The cathespins have
also expanded in rodents, with six showing high expression
(\Sexpr{gene.expression("mus_musculus","Ctsj",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Cts6",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Ctsr",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Ctsq",complete=TRUE)},
\Sexpr{gene.expression("mus_musculus","Cts3",complete=TRUE)}, and
\Sexpr{gene.expression("mus_musculus","Ctsm",complete=TRUE)},
Supplemental \cref{tab:all_mus_musculus_placenta,fig:pag2_prss16_ctsl_psg_evolution_tree}).
Finally, \textit{Peg10}, an imprinted gene required for proper
formation of the placenta in mice is also highly expressed
(\Sexpr{gene.expression("mus_musculus","Peg10")}), as previously
reported \cite{Ono.ea2006:DeletionofPeg10imprinted}.

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \galili are shown in
Supplemental \cref{tab:all_n_galili_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_spalax_galili_placenta}.
Components of annexin complexes are highly expressed
(\Sexpr{gene.expression("spalax_galili","S100a6",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","S100a11",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","S100a10",complete=TRUE)},
and
\Sexpr{gene.expression("spalax_galili","Anxa2",complete=TRUE)}),
as are members of the thymosin family
(\Sexpr{gene.expression("spalax_galili","Tmsb4x",complete=TRUE)}
and
\Sexpr{gene.expression("spalax_galili","Tmsb10",complete=TRUE)}),
various cytoskeleton related proteins
(\Sexpr{gene.expression("spalax_galili","Actc1",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","Actg2",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","Acta2",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","Myl9",complete=TRUE)},
\Sexpr{gene.expression("spalax_galili","Vim",complete=TRUE)}), and
transmembrane protein of currently unknown function, 
\Sexpr{gene.expression("spalax_galili","Gm10134",complete=TRUE)}.

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \carmeli are shown in
Supplemental \cref{tab:all_s_carmeli_placenta} with a list of all highly
expressed genes in Supplemental
\cref{tab:all_spalax_carmeli_placenta}. Numerous hemoglobin subunits
ware highly expressed
(\Sexpr{gene.expression("spalax_carmeli","Hbb-bh1",complete=TRUE)},
\Sexpr{gene.expression("spalax_carmeli","Hba-a1",complete=TRUE)}, and
\Sexpr{gene.expression("spalax_carmeli","Hbb-bs",complete=TRUE)}), as
well as the relaxin
\Sexpr{gene.expression("spalax_carmeli","Rln1",complete=TRUE)},
multiple cathespins
(\Sexpr{gene.expression("spalax_carmeli","Ctsm",complete=TRUE)},
\Sexpr{gene.expression("spalax_carmeli","BC051665",complete=TRUE)},
and \Sexpr{gene.expression("spalax_carmeli","Ctsl",complete=TRUE)},
\cref{fig:pag2_prss16_ctsl_psg_evolution_tree}~B), and many prolactin-related genes
(\Sexpr{gene.expression("spalax_carmeli","Prl4a1",complete=TRUE)},
\Sexpr{gene.expression("spalax_carmeli","Prl6a1",complete=TRUE)},
\Sexpr{gene.expression("spalax_carmeli","Prl3b1",complete=TRUE)}, and
\Sexpr{gene.expression("spalax_carmeli","Prl7c1",complete=TRUE)},
\cref{fig:prolactin_evolution_tree}) which shows a similar pattern of
expansion seen in \mouse.

@@

The top 50 non-housekeeping expressed genes as measured by \ac{FPKM}
in the term placenta of \dog are shown in
Supplemental \cref{tab:all_canis_familiaris_placenta} with a list of all genes in Supplemental
\cref{tab:all_canis_familiaris_placenta}. The gene with the highest
expression %
(\Sexpr{gene.expression("canis familiaris","TMSB10")}) is the thymosin 
\textit{TMSB10} %
(\Sexpr{gene.expression("canis familiaris","ENSCAFG00000002115")}) %
is the hormone relaxin (\textit{RLN2}). Other highly expressed genes
among the top 50 (Supplemental \cref{tab:all_canis_familiaris_placenta}) include genes
involved in the cytoskeleton (\textit{TMSB10}, \textit{TMSB4X}),
anticoagulation (\textit{TFPI2}, \textit{PLA2G7}), innate immunity
(\textit{LCN2}, \textit{GM2A}), lipid metabolism (\textit{APOE}),
inhibition of matrix metalloproteinases (\textit{TIMP2},
\textit{TIMP3}), and growth factors (\textit{PGF}, \textit{EGFR},
\textit{HTRA1}). The gene \textit{LGALS1}, which shows evidence of
immunosupression during pregnancy
\cite{Than.ea2008:Emergenceofhormonaland}, is highly expressed %
(\Sexpr{gene.expression("canis familiaris","LGALS1")}). %
In addition, an ortholog of \textit{SSLP1} (ENSCAFG00000010638) a gene
expressed in the placentas of \cow (%
\Sexpr{gene.expression("bos taurus","SSLP1",complete=TRUE)}, %
Supplemental \cref{tab:all_bos_taurus_placenta}) and \sheep
(\Sexpr{gene.expression("ovis aries","SOLD1",complete=TRUE)},
Supplemental \cref{tab:all_sheep_placenta}) but is not present in primates
\cite{Ushizawa.ea2009:Characterizationandexpressionanalysis}
(\cref{fig:sslp1_evolution_tree}), is also expressed at high levels
%
(\Sexpr{gene.expression("canis familiaris","ENSCAFG00000010638")}). %
There is no placental expression of the estrogen synthase gene
(\textit{CYP19A1}) in \dog 
(\Sexpr{gene.expression("canis familiaris","CYP19A1")}
\cref{fig:estrogen_evolution_tree}), 
and expression of \textit{HSD3B7} is also low
(\Sexpr{gene.expression("canis familiaris","HSD3B7")}, 
\cref{fig:progesterone_evolution_tree}).

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \textit{Sus scrofa} are shown in
Supplemental \cref{tab:all_pig_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_sus_scrofa_placenta}. The top two
expressed transcripts,
\Sexpr{gene.expression("sus_scrofa","ENSSSCG00000025162",complete=TRUE)}
and
\Sexpr{gene.expression("sus_scrofa","ENSSSCG00000021356",complete=TRUE)},
are miRNA which correspond to the \cow miRNA
\href{http://www.mirbase.org/cgi-bin/mirna_entry.pl?acc=MI0013080}{bta-mir-2904-1}
and
\href{http://www.mirbase.org/cgi-bin/mirna_entry.pl?acc=MIMAT0013845}{bta-mir-2887-1},
respectively, both of which are also expressed in \cow ovary
\cite{Hossain.ea2009:Identificationandcharacterizationof}, although
neither of which have an obvious target gene on the basis of
\textit{in silico} analysis
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget}. Other highly
expressed genes include members of the insulin growth factor family
(\Sexpr{gene.expression("sus_scrofa","IGF2",complete=TRUE)} and
\Sexpr{gene.expression("sus_scrofa","IGFBP3",complete=TRUE)}), members
of annexin complexes
(\Sexpr{gene.expression("sus_scrofa","S100A1",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","ANXA2",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","S100A12",complete=TRUE)}, and
\Sexpr{gene.expression("sus_scrofa","S100A10",complete=TRUE)}), the
galectin \Sexpr{gene.expression("sus_scrofa","LGALS3",complete=TRUE)}
and multiple cytoskeleton and extracellular matrix proteins
(\Sexpr{gene.expression("sus_scrofa","KRT19",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","COL1A1",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","BETA-ACTIN",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","KRT9",complete=TRUE)},
\Sexpr{gene.expression("sus_scrofa","CLDN4",complete=TRUE)}, and
\Sexpr{gene.expression("sus_scrofa","KRT9",complete=TRUE)}).

@@

The 50 non-housekeeping genes with highest expression (by \ac{FPKM})
in the term placenta of \horse are shown in
Supplemental \cref{tab:all_equus_caballus_placenta} with a list of all highly expressed
genes in Supplemental \cref{tab:all_equus_caballus_placenta}. The gene
with the highest expression was the relaxin,
\Sexpr{gene.expression("equus_caballus","RLN",complete=TRUE)},
followed by two miRNA
(\Sexpr{gene.expression("equus_caballus","ENSECAG00000025328",complete=TRUE)}
and
\Sexpr{gene.expression("equus_caballus","ENSECAG00000027056",complete=TRUE)})
with no known \textit{in silico} targets
\cite{Agarwal.ea2015:PredictingeffectivemicroRNAtarget}, the thymosin
\Sexpr{gene.expression("equus_caballus","TMSB4",complete=TRUE)},
members of the insulin growth factor family
(\Sexpr{gene.expression("equus_caballus","IGF2",complete=TRUE)} and
\Sexpr{gene.expression("equus_caballus","IGFBP3",complete=TRUE)}) and
multiple members of annexin complexes
(\Sexpr{gene.expression("equus_caballus","S100A6",complete=TRUE)},
\Sexpr{gene.expression("equus_caballus","S100A1",complete=TRUE)}, and
\Sexpr{gene.expression("equus_caballus","ANXA2",complete=TRUE)}),
similar to the pattern of expression seen in \pig and \dog.
